---
title: "Bat virus underroost shedding model"
author: "Benny Borremans"
output:
  pdf_document:
    toc: yes
  word_document:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---

```{r global_options, include = FALSE,results='hide'}
knitr::opts_chunk$set(echo = FALSE, results = 'hide',warning = FALSE, message = FALSE,tidy.opts=list(width.cutoff=90),fig.height=4,fig.width=4,dev='pdf',fig.align='center',out.width = "70%")

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))  # set file folder as working directory, need to save file first

library(tidyverse) # data handling
library(ggplot2) # plotting
library(patchwork) # combine plots
library(RColorBrewer) # use color palette brewer
library(lubridate)   # date manipulations
library(MASS) # fit distributions 
library(brms)   # bayesian regression models with stan, through brms   
library(rstan)   # bayesian regression models with stan, through rstan


# set colors for plots 
col.no.groups = brewer.pal(n = 11,"RdYlBu")[4]
col.2.groups = brewer.pal(n = 11,"RdYlBu")[c(1,10)]
col.4.groups = brewer.pal(n = 11,"RdYlBu")[c(1,4,8,11)]




```

\vspace{20pt}

[bennyborremans@bbresearch.org](mailto:bennyborremans@bbresearch.org){.email}

\vspace{10pt}

Last update: `r format(Sys.Date(),"%d %b %Y")`.

\vspace{10pt}

The table of contents can be clicked to jump straight to specific sections.


\pagebreak   

Goal = create a model of underroost bat virus shedding.    

Sheets placed below roosts collect urine from an estimated number of bats.   
Urine samples on the sheet are pooled, and tested for RNA concentration using RT-PCR.   
Total sample volume depend on the number and volume of urine droplets on the sheets.    
The number and species of bats above the sheet are estimated, but not all bats can always be observed, and bats can move after/before observation.   
Samples are stored in one of two buffers (AVL/VTM), or without buffer (NB).   
Buffer type affects PCR sensitivity.    

The end result (Ct value in a pooled sample) depends on all these factors,    
which makes it difficult to estimate shedding prevalence in the population.   

The goal of this project is two-fold:    
(1) Create a simulation model of the different processes that are believed to be involved, to get better insights and build intuition.    
(2) Create a model to estimate shedding prevalence in the population from the available data, capturing as much of the observation process as reasonably possible.    



# Simulation model     


\pagebreak


## Model number of bats above a sheet   


```{r}
# import data from database:
# library(PREEMPTdatabase)
# dbcon = preempt_database_connect("bennyborremans_preempt_creds.txt")
# b0 = australia_underroost(dbcon)
# saveRDS(b0,"australia_underroost_downloaded_20220607.RDS")

b0 = readRDS("australia_underroost_downloaded_20220607.RDS")

# explanation of columns:
# preempt_data_dictionary()
b0.dict = read.csv("preempt_database_data_dictionary_australia_underroost.csv",stringsAsFactors = F)


# extract site from accession id (the site column is mostly empty so overwriting)
b0$site = substr(b0$accession_update,start = 3,stop = 5)

b0$date = as.Date(b0$date)

b0$urine_vol_num = as.numeric(b0$urine_vol)
b0$urine_avl_vol_1_num = as.numeric(b0$urine_avl_vol)
b0$urine_avl_vol_2_num = as.numeric(b0$urine_avl_vol_2)
b0$urine_vtm_vol_1_num = as.numeric(b0$urine_vtm_vol_1)
b0$urine_vtm_vol_2_num = as.numeric(b0$urine_vtm_vol_2)

b0$evap[which(b0$evap %in% c("NA","NR","NC","UK"))] = NA
b0$evap[which(b0$evap == "pevap")] = "partial"
b0$evap[which(b0$evap == "evap")] = "full"

b0$season = NA
b0$season[which(month(b0$date) %in% c(12,1,2))] = "summer"
b0$season[which(month(b0$date) %in% c(3,4,5))] = "autumn"
b0$season[which(month(b0$date) %in% c(6,7,8))] = "winter"
b0$season[which(month(b0$date) %in% c(9,10,11))] = "spring"
table(b0$season)

```

\vspace{50pt}   


There are counts for 2 species, black flying foxes and grey-headed flying foxes.    
Counts can be morning, afternoon, and/or overall.    
Observations have a level of confidence in the count and/or species.   
These confidence levels are available for bff only (\textcolor{blue}{do they cover both bff and ghff, or bff only?}).    


\vspace{10pt}   

Using only observations with a high level of confidence.    
Using only morning observations (as these seem to be the most common).    
Removing observations that are not exact (e.g. "5+").    
All sites pooled.     

\vspace{10pt}   

```{r}
b1 = b0 %>%
       mutate(bff_conf.num = as.numeric(bff_conf)) %>%  # change to numeric
       filter(bff_conf.num == 1) %>%    # only keep confidence level 1
       mutate(bffm.count = as.numeric(bffm)) %>%    # change to numeric, so entries like "5+" become NA
       filter(is.finite(bffm.count)) %>%    # only keep non-NA
       mutate(ghffm.count = as.numeric(ghffm)) %>%   # change to numeric
       mutate(total.count = bffm.count + ghffm.count)   # sum bff and ghff counts



```

\vspace{20pt}    

Histograms of counts:    

\vspace{20pt}    

```{r fig.width = 15, fig.height = 5, out.width = "90%"}

plot1 = ggplot(data = b1, aes(bffm.count)) +
       geom_histogram(binwidth = 1, fill = col.2.groups[1], color = "black") +
       ggtitle("bff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))

plot2 = ggplot(data = b1, aes(ghffm.count)) +
       geom_histogram(binwidth = 1, fill = col.2.groups[2], color = "black") +
       ggtitle("ghff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))

plot3 = ggplot(data = b1, aes(total.count)) +
       geom_histogram(binwidth = 1, fill = col.no.groups, color = "black") +
       ggtitle("bff + ghff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))


plot1 + plot2 + plot3

```


\vspace{20pt}   


Are higher numbers actually more rare, or just more uncertain?    

==> check lower confidence counts.     


```{r}
b2 = b0 %>%
       mutate(bff_conf.num = as.numeric(bff_conf)) %>%  # change to numeric
       filter(bff_conf %in% c(0,0.5,2)) %>%    # only keep confidence level 1
       mutate(bffm.count = as.numeric(bffm)) %>%    # change to numeric, so entries like "5+" become NA
       filter(is.finite(bffm.count)) %>%    # only keep non-NA
       mutate(ghffm.count = as.numeric(ghffm)) %>%   # change to numeric
       mutate(total.count = bffm.count + ghffm.count)   # sum bff and ghff counts


```

\vspace{20pt}    

Histograms of counts:    

\vspace{20pt}    

```{r fig.width = 15, fig.height = 5, out.width = "90%"}

plot1 = ggplot(data = b2, aes(bffm.count)) +
       geom_histogram(binwidth = 1, fill = col.2.groups[1], color = "black") +
       ggtitle("bff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))

plot2 = ggplot(data = b2, aes(ghffm.count)) +
       geom_histogram(binwidth = 1, fill = col.2.groups[2], color = "black") +
       ggtitle("ghff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))

plot3 = ggplot(data = b2, aes(total.count)) +
       geom_histogram(binwidth = 1, fill = col.no.groups, color = "black") +
       ggtitle("bff + ghff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))


plot1 + plot2 + plot3

```

\vspace{20pt}   

Lower-confidence counts are not higher, except for 2 ghff counts.    




\vspace{20pt}   


Are higher numbers actually more rare, or just written down as N+?    


```{r}

b4 = b0[grep("+", b0$bffm, fixed = T),]

```

Not likely, there are only `r nrow(b4)` entries with a + sign,  
and these are one of: `r unique(b4$bffm)`.    


\pagebreak 

Are there differences between sites?    


\vspace{20pt}   

Using only sites with more than 100 observations.    


```{r}
site.freq = as.data.frame(table(b1$site))
site.over.100 = droplevels(site.freq$Var1[which(site.freq$Freq>100)])
```

```{r fig.width = 15, fig.height = 10, out.width = "90%"}

ggplot(data = b1[which(b1$site %in% site.over.100),], aes(bffm.count)) +
       geom_histogram(binwidth = 1, fill = col.2.groups[1], color = "black") +
       ggtitle("bff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm")) +
       facet_wrap(~ site)


ggplot(data = b1[which(b1$site %in% site.over.100),], aes(ghffm.count)) +
       geom_histogram(binwidth = 1, fill = col.2.groups[2], color = "black") +
       ggtitle("ghff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm")) +
       facet_wrap(~ site)


```


\vspace{20pt}


==> all look very similar, except many more 0 counts at CLU.   
\textcolor{blue}{Any reason for this? Different conditions?}       
I didn't find anything in the notes, and the "bats" column mostly says "stable".    


```{r}
# View(b1[which(b1$site=="CLU"),])
```


\vspace{50pt}   

Any difference between seasons?    

Histograms for different seasons.    
bff only, most data available, don't need figure overload.    

```{r fig.width = 14, fig.height = 8, out.width = "60%"}
ggplot(data = b1, aes(bffm.count)) +
       geom_histogram(binwidth = 1, fill = col.2.groups[1], color = "black") +
       ggtitle("bff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm")) +
       facet_wrap(~ season)
```

\vspace{20pt}   

There seems to be an effect of season, that probably should be taken into account.    

While spring and summer look like Poisson distributions,  
autumn and winter seem closer to a mixture of a Bernoulli and a Poisson (probability of seeing any bats + if there are bats, how many).     

==> Fit model to positive counts only, as simulating 0 bats will not be useful.    

```{r}
b1 = b1 %>%
       filter(bffm.count > 0)
```


\vspace{50pt}   

Try Poisson distribution    

\vspace{20pt} 

All bff data pooled, across seasons and sites:        

```{r}
fit.pois1 = fitdistr(b1$bffm.count, densfun = "poisson")
```


Lambda = `r as.numeric(fit.pois1$estimate)`.    

Fitted distribution:    

```{r fig.width = 10, fig.height = 7, out.width = "40%"}

fit.plot = data.frame(x = 1:20,
                      y = dpois(x = 1:20, fit.pois1$estimate))

ggplot(data = b1, aes(bffm.count)) +
       geom_histogram(aes(y = ..density..), binwidth = 1, fill = col.2.groups[1], color = "black", alpha = 1) +
       geom_line(data = fit.plot, aes(x = x, y = y), color = "black", size = 1) +
       ggtitle("bff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))
```

This distribution fits pretty well for the lower counts, but does not allow for the occasional higher numbers.    

==> try a distribution with some more variance: negative binomial.    


All bff data pooled, across seasons and sites:        

```{r}
fit.negbin1 = fitdistr(b1$bffm.count, densfun = "negative binomial")
```


Fitted distribution:    

```{r fig.width = 10, fig.height = 7, out.width = "40%"}

fit.plot = data.frame(x = 1:20,
                      y = dnbinom(x = 1:20, size =  fit.negbin1$estimate[1], mu = fit.negbin1$estimate[2]))

ggplot(data = b1, aes(bffm.count)) +
       geom_histogram(aes(y = ..density..), binwidth = 1, fill = col.2.groups[1], color = "black", alpha = 1) +
       geom_line(data = fit.plot, aes(x = x, y = y), color = "black", size = 1) +
       ggtitle("bff") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))
```



==> almost no difference, but a marginally wider tail for the negative binomial distribution.    

A distribution with a heavier tail would be a bit better, but for the simulation we can use the negative binomial distribution.    



\vspace{20pt} 

All bff data pooled, per season, across all sites:        

```{r}
fit.negbin1.spring = fitdistr(b1$bffm.count[which(b1$season=="spring")], densfun = "negative binomial")
fit.negbin1.summer = fitdistr(b1$bffm.count[which(b1$season=="summer")], densfun = "negative binomial")
fit.negbin1.autumn = fitdistr(b1$bffm.count[which(b1$season=="autumn")], densfun = "negative binomial")
fit.negbin1.winter = fitdistr(b1$bffm.count[which(b1$season=="winter")], densfun = "negative binomial")

fit.negbin1.seasons = data.frame(season = c("spring","summer","autumn","winter"),
                               size = c(fit.negbin1.spring$estimate[1], fit.negbin1.summer$estimate[1], fit.negbin1.autumn$estimate[1], fit.negbin1.winter$estimate[1]),
                               mu = c(fit.negbin1.spring$estimate[2], fit.negbin1.summer$estimate[2], fit.negbin1.autumn$estimate[2], fit.negbin1.winter$estimate[2]))
```



Fitted distributions:    

```{r fig.width = 10, fig.height = 7, out.width = "60%"}

fit.plot = data.frame(x = 1:20,
                      y = dnbinom(x = 1:20, size = fit.negbin1.spring$estimate[1], mu = fit.negbin1.spring$estimate[2]))

plot1 = ggplot(data = b1[which(b1$season=="spring"),], aes(bffm.count)) +
       geom_histogram(aes(y = ..density..), binwidth = 1, fill = col.2.groups[1], color = "black", alpha = 1) +
       geom_line(data = fit.plot, aes(x = x, y = y), color = "black", size = 2) +
       ggtitle("bff - spring") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))

fit.plot = data.frame(x = 1:20,
                      y = dnbinom(x = 1:20, size = fit.negbin1.summer$estimate[1], mu = fit.negbin1.summer$estimate[2]))

plot2 = ggplot(data = b1[which(b1$season=="summer"),], aes(bffm.count)) +
       geom_histogram(aes(y = ..density..), binwidth = 1, fill = col.2.groups[1], color = "black", alpha = 1) +
       geom_line(data = fit.plot, aes(x = x, y = y), color = "black", size = 2) +
       ggtitle("bff - summer") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))

fit.plot = data.frame(x = 1:20,
                      y = dnbinom(x = 1:20, size = fit.negbin1.autumn$estimate[1], mu = fit.negbin1.autumn$estimate[2]))

plot3 = ggplot(data = b1[which(b1$season=="autumn"),], aes(bffm.count)) +
       geom_histogram(aes(y = ..density..), binwidth = 1, fill = col.2.groups[1], color = "black", alpha = 1) +
       geom_line(data = fit.plot, aes(x = x, y = y), color = "black", size = 2) +
       ggtitle("bff - autumn") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))

fit.plot = data.frame(x = 1:20,
                      y = dnbinom(x = 1:20, size = fit.negbin1.winter$estimate[1], mu = fit.negbin1.winter$estimate[2]))

plot4 = ggplot(data = b1[which(b1$season=="winter"),], aes(bffm.count)) +
       geom_histogram(aes(y = ..density..), binwidth = 1, fill = col.2.groups[1], color = "black", alpha = 1) +
       geom_line(data = fit.plot, aes(x = x, y = y), color = "black", size = 2) +
       ggtitle("bff - winter") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))

(plot1 + plot2) /
       (plot3 + plot4)
```


All combined:    

```{r fig.width = 11, fig.height = 6}
fit.plot = data.frame(x = rep(1:20,5),
                      y = c(dnbinom(x = 1:20, size = fit.negbin1$estimate[1], mu = fit.negbin1$estimate[2]),
                            dnbinom(x = 1:20, size = fit.negbin1.spring$estimate[1], mu = fit.negbin1.spring$estimate[2]),
                            dnbinom(x = 1:20, size = fit.negbin1.summer$estimate[1], mu = fit.negbin1.summer$estimate[2]),
                            dnbinom(x = 1:20, size = fit.negbin1.autumn$estimate[1], mu = fit.negbin1.autumn$estimate[2]),
                            dnbinom(x = 1:20, size = fit.negbin1.winter$estimate[1], mu = fit.negbin1.winter$estimate[2])),
                      season = rep(c("all","spring", "summer", "autumn", "winter"), each = 20))

ggplot(data = fit.plot, aes(x = x, y = y, color = season)) +
       geom_line(size = 1) +
       scale_color_manual(values = c("black",brewer.pal(n = 11,"RdYlBu")[c(1,3,5,10)]), name = "") +
       ggtitle("bff") +
       xlab("Count") +
       ylab("Probability density") +
       scale_x_continuous(breaks = 0:50) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))
```

\vspace{20pt}   

==> fitted negative binomial distributions are very similar across seasons.    



**Conclusion:**     
Bat count can be modeled adequately using a Negative Binomial distribution, with:    
$N_{bats} \sim NegBinom(size = `r round(fit.negbin1$estimate[1],1)`, mu = `r round(fit.negbin1$estimate[2],1)`)$   



\pagebreak




## Model sample volume    


How are sample volumne and the number of bats related?    
More bats = more urine, but how strong is this correlation, and what is its shape?    

Considerations that need to be made when looking at this:    
- There can be evaporation.    
- The sample is added to buffer (how is this recorded in the data?)    



\vspace{50pt} 

How much urine can different numbers of bats produce?    
What is the variation in collected urine volume, given a certain bat count?    

And for each bat count, how does evaporation affect the distribution of volumes?    

How common is evaporation in each season?    

(using only confidence level = 1 data)    
There are some volumes indicated as <X (e.g. < 140), which could be included when modeling by allowing censoring,    
but to keep things simple these are just removed.    

\vspace{20pt} 

```{r results = 'hide', fig.width = 10, fig.height = 8, out.width = "40%"}
# urine is divided into multiple collection tubes,   
# so we need to sum that first.

table(b1$urine_vol_num)
table(b1$urine_avl_vol_1_num)
table(b1$urine_avl_vol_2_num)   # all 0
table(b1$urine_vtm_vol_1_num)
table(b1$urine_vtm_vol_2_num)

b5 = b0 %>%
       mutate(bff_conf.num = as.numeric(bff_conf)) %>%  # change to numeric
       filter(bff_conf.num == 1) %>%    # only keep confidence level 1
       mutate(bffm.count = as.numeric(bffm)) %>%    # change to numeric, so entries like "5+" become NA
       filter(is.finite(bffm.count)) %>%    # only keep non-NA
       mutate(ghffm.count = as.numeric(ghffm)) %>%   # change to numeric
       mutate(total.count = bffm.count + ghffm.count) %>%  # sum bff and ghff counts
       rowwise() %>%
       mutate(urine_vol_total = sum(c(urine_vol_num,urine_avl_vol_1_num,urine_avl_vol_2_num,urine_vtm_vol_1_num,urine_vtm_vol_2_num),na.rm=T))

# View(b5[which(b5$urine_vol_total==1000),])

ggplot(data = b5, aes(urine_vol_total)) +
       geom_histogram(fill = col.2.groups[1], color = "black") +
       ggtitle("Total urine volume collected") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = seq(0,3000, by = 250)) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))

```


\vspace{20pt}   

==> there are many zeros,    
and a lot of 1000s.    
Those 1000s are mostly the result of 500 for each of the vtm tubes.    
\textcolor{blue}{Does this mean they weren't the maximum volume possible?}     

Removing these samples for now:    


```{r fig.width = 10, fig.height = 8, out.width = "40%"}
b6 = b5[-which(b5$urine_vtm_vol_1_num==500 & b5$urine_vtm_vol_2_num==500),]

ggplot(data = b6, aes(urine_vol_total)) +
       geom_histogram(fill = col.2.groups[1], color = "black") +
       ggtitle("Total urine volume collected") +
       xlab("Observed count") +
       ylab("Frequency count") +
       scale_x_continuous(breaks = seq(0,3000, by = 250)) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))

```

\vspace{50pt}   

```{r fig.width = 10, fig.height = 8, out.width = "40%"}

ggplot(data = b6[-which(is.na(b6$evap)),], aes(urine_vol_total)) +
       geom_histogram(fill = col.2.groups[1], color = "black") +
       ggtitle("Total volume collected vs evaporation status") +
       xlab("Volume") +
       ylab("Frequency count") +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm")) +
       facet_wrap(~evap)
```


\vspace{50pt}   

Correlation between bat counts and total volume.    
Excluding counts of 0.    
And excluding "full" evaporation, as that is always 0.    

\textcolor{blue}{Even then, many total volumes are still 0, why is that?}    
Excluding total volume = 0 data for now.    


```{r}
b7 = b6[-which(b6$bffm.count==0),] %>%
       filter(!(is.na(bffm.count) | is.na(urine_vol_total) | is.na(evap))) %>%
       filter(evap %in% c("none","partial")) %>%
       filter(urine_vol_total > 0)
```

```{r fig.width = 8, fig.height = 7, out.width = "50%"}
plot1 = ggplot(data = b7, aes(x = bffm.count, y = urine_vol_total)) +
       geom_point(color = col.2.groups[1]) +
       xlab("bff count") +
       ylab("total urine volume") +
       scale_x_continuous(breaks = seq(1,20,by = 2)) +
       scale_y_continuous(breaks = seq(0,4000,by = 250)) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))
plot1
```

\vspace{20pt}    

Regression model:    

total urine volume ~ bat count + evaporation status     



```{r}
fit1.brm = brm(urine_vol_total ~ bffm.count + evap, data = b7, family = Gamma(link='log'))

fit1.brm.posteriors = as_draws_df(fit1.brm)
```

Output (log):    

```{r results = 'markup', fig.width = 10, fig.height = 15, out.width = "60%"}
plot(fit1.brm)

sum.fit1.brm = summary(fit1.brm)
```

Back-transformed (exp) coefficients:    

```{r results = 'markup'}
exp(sum.fit1.brm$fixed[,1])
```


Fitted functions/coefficients:   

```{r fig.height = 6, fig.width = 8, out.width = "40%"}
plot(conditional_effects(fit1.brm, effects = 'bffm.count'), points = T)
plot(conditional_effects(fit1.brm, effects = 'evap'), points = T)

```


There is a positive correlation between bat count and total urine volume.    
There is a minor effect of evaporation.    


\vspace{20pt}  

Can we use this model to adequately simulate urine volumes, given a bat count and evaporation status?    

Bat counts are simulated using the negative binomial distribution fitted above (excluding 0s).    
Evaporation status is simulated by randomly choosing 'none' or 'partial'.   
Urine volume is then simulated using a gamma distribution with parameter values randomly selected from the model posterior.    

Red = predicted, blue = observed.    


```{r fig.width = 8, fig.height = 6, out.width = "60%"}

iterations = 1000

sim.dat = data.frame(bat.count = rnbinom(size = fit.negbin1$estimate[1], mu = fit.negbin1$estimate[2], n = iterations),
                     evap = sample(c(0,1),size = iterations, replace = T),  # 0 = none, 1 = partial
                     pred.volume = NA) %>%
       filter(bat.count > 0)


for(i in 1:nrow(sim.dat)){
       random.iteration = sample(1:nrow(fit1.brm.posteriors),1)
       shape.par = fit1.brm.posteriors$shape[random.iteration]
       scale.par = exp(fit1.brm.posteriors$b_Intercept[random.iteration] + fit1.brm.posteriors$b_bffm.count[random.iteration] * sim.dat$bat.count[i] + fit1.brm.posteriors$b_evappartial[random.iteration] * sim.dat$evap[i])
       sim.dat$pred.volume[i] = rgamma(n = 1, shape = shape.par, scale = scale.par)
}


ggplot(data = sim.dat, aes(x = bat.count, y = pred.volume)) +
       geom_point(color = col.2.groups[1], alpha = 0.5) +
       geom_point(data = b7, aes(x = bffm.count + 0.3, y = urine_vol_total), col = col.2.groups[2], alpha = 0.5) +
       xlab("bff count") +
       ylab("total urine volume") +
       scale_x_continuous(breaks = seq(1,20,by = 2)) +
       #scale_y_continuous(breaks = seq(0,4000,by = 250), limits = c(0,4000)) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))


```


\vspace{20pt}    

Simulation output is decently similar, but the large standard deviation results in the prediction of urine volumes larger than the ones observed.   
==> should be excluded:   

\vspace{20pt}  

```{r fig.width = 8, fig.height = 6, out.width = "60%"}

iterations = 1000

sim.dat = data.frame(bat.count = rnbinom(size = fit.negbin1$estimate[1], mu = fit.negbin1$estimate[2], n = iterations),
                     evap = sample(c(0,1),size = iterations, replace = T),  # 0 = none, 1 = partial
                     pred.volume = NA) %>%
       filter(bat.count > 0)


for(i in 1:nrow(sim.dat)){
       random.iteration = sample(1:nrow(fit1.brm.posteriors),1)
       shape.par = fit1.brm.posteriors$shape[random.iteration]
       scale.par = exp(fit1.brm.posteriors$b_Intercept[random.iteration] + fit1.brm.posteriors$b_bffm.count[random.iteration] * sim.dat$bat.count[i] + fit1.brm.posteriors$b_evappartial[random.iteration] * sim.dat$evap[i])
       sim.dat$pred.volume[i] = rgamma(n = 1, shape = shape.par, scale = scale.par)
}

sim.dat = sim.dat[-which(sim.dat$pred.volume > max(b7$urine_vol_total)),]

ggplot(data = sim.dat, aes(x = bat.count, y = pred.volume)) +
       geom_point(color = col.2.groups[1], alpha = 0.5) +
       geom_point(data = b7, aes(x = bffm.count + 0.3, y = urine_vol_total), col = col.2.groups[2], alpha = 0.5) +
       xlab("bff count") +
       ylab("total urine volume") +
       scale_x_continuous(breaks = seq(1,20,by = 2)) +
       #scale_y_continuous(breaks = seq(0,4000,by = 250), limits = c(0,4000)) +
       theme_light() +
       theme(plot.title  =  element_text(size = 11),
             text = element_text(size = 11),
             axis.text = element_text(size = 11),
             axis.title.y  =  element_text(margin  =  ggplot2::margin(r = 10)),
             plot.margin = unit(c(0.5,0.75,0.5,0.5),"cm"))


```



\pagebreak    

## Shedding and Ct values    


What is the distribution of Ct values being shed by individuals?    
How is this affected by buffer type?    
How does this translate to genome copies / virions?    
What is the distribution of Ct values being in underroost samples?    






# Prevalence estimation model    

```{r}

```





